---
title: "PCA for SynNiche"
author: "TRS"
date: "11/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This code provides the framework to construct a PCA of multiple Syndiniales species and determine the variables that describe observed differences in community composition. 
First, import a dataframe with the species observations and associated environmental factors.
```{r}
occ.sp = read.table("species_forECOSPAT.txt", header=T, row.names =1, sep ="\t")
head(occ.sp)
```

To construct a PCA with many species, use the following code:

First, group by "species" or strain of Syndiniales. Here we are only using presence data.

```{r}
spGroup <- factor(occ.sp$species)
dev.off() ##clear plot
```

Run PCA
This is the basic plot, which does not show variable vectors

```{r}
synPCA <- prcomp(scale(occ.sp[,-1]))
plot(synPCA$x[,1:2], col = spGroup)##Basic plot
```

For more advanced plotting, use:

```{r}
library(pcaMethods)
synPCAmethods <- pca(occ.sp[,-1], scale = "uv", center = T, nPcs = 2, method = "svd")
dev.off()
par(mar=c(1,1,1,1)) 
slplot(synPCAmethods, scoresLoadings = c(T,T), sl = spGroup)
summary(synPCAmethods)
```

For publications, I would use the following toolbox. This one draws ellipses to show each "species niche" and plots the variable vectors.

```{r}
library(factoextra)
library(ggplot2)
fviz_pca_biplot(synPCA, axes=c(1, 2), geom.ind="point", geom.var= "arrow","text", col.ind = spGroup,addEllipses = TRUE, ellipse.type = "confidence", legend.title = "Groups", repel = FALSE)
```

Even this method does not label the variables, but that information can be found on the previous plot.